Simulation

### 2023-10-31 simulation.qmd
Wrap up univariate brms

```{r}
# model <- cmdstan_model("model-ar1.stan", compile = FALSE)
# model$format(
#   canonicalize = list("deprecations"), 
#   overwrite_file = TRUE, 
#   backup = FALSE
# )
# model$compile()
model <- cmdstan_model(exe_file = "model-ar1")
path <- "models/cmdstanr-example.rds"
if (!file.exists(path)) {
post <- model$sample(
  data = dat, 
  adapt_delta = 0.99
) |> 
  as_draws_df()
saveRDS(post, path)
} else {post <- readRDS(path)}
```

@@

```{r}
post <- post |> 
  select(starts_with("beta_"), ends_with("_raw")) |> 
  rename_with(~str_remove(., "_raw"))
post |> 
  pivot_longer(everything()) |> 
  ggplot(aes(value, name)) +
  scale_x_continuous(
    "Parameter value"
  ) +
  stat_histinterval(
    normalize = "xy",
    scale = 0.25, 
    size = 1
  ) +
  geom_point(
    data = as_tibble(dat) |> 
      select(sigma:beta_scale) |> 
      distinct() |> 
      pivot_longer(everything()),
    position = position_nudge(y = -0.1),
    color = "red"
  ) +
  theme(
    axis.title.y = element_blank()
  )
```

@@

```{r}
dat <- as_tibble(dat)
dat <- dat |> 
  mutate(yt1 = lag(y), .by = individual)
dat
```

>>>>

```{r}
# model <- cmdstan_model("model-ar1.stan", compile = FALSE)
# model$format(
#   canonicalize = list("deprecations"), 
#   overwrite_file = TRUE, 
#   backup = FALSE
# )
# model$compile()
model <- cmdstan_model(exe_file = "model-ar1")
path <- "models/cmdstanr-example.rds"
if (!file.exists(path)) {
post <- model$sample(
  data = dat, 
  adapt_delta = 0.99
) |> 
  as_draws_df()
saveRDS(post, path)
} else {post_cmdstanr <- readRDS(path)}
```

@@

```{r}
post_cmdstanr <- post_cmdstanr |> 
  select(starts_with("beta_"), ends_with("_raw")) |> 
  rename_with(~str_remove(., "_raw"))
post_cmdstanr |> 
  pivot_longer(everything()) |> 
  ggplot(aes(value, name)) +
  scale_x_continuous(
    "Parameter value"
  ) +
  stat_histinterval(
    normalize = "xy",
    scale = 0.25, 
    size = 1
  ) +
  geom_point(
    data = as_tibble(dat) |> 
      select(sigma:beta_scale) |> 
      distinct() |> 
      pivot_longer(everything()),
    position = position_nudge(y = -0.1),
    color = "red"
  ) +
  theme(
    axis.title.y = element_blank()
  )
```

@@

```{r}
dat <- as_tibble(dat)
dat <- dat |> 
  mutate(yt1 = lag(y), .by = individual)
head(dat)
```

We specify the model as a missing data model as follows

```{r}
bform <- bf(y | mi() ~ -1 + mi(mean) + b2*(yt1 - mi(mean))) + 
  bf(mean | mi() ~ 1 + (1 | individual))
fit <- brm(
  bform,
  data = dat
)
```

However this needs some brms syntax edits. We also change the parameter names to fit the blog post

```{r}
bform <- bf(
  y | mi() ~ 0 + mi(alpha) + beta * (yt1 - mi(alpha)),
  alpha | mi() ~ 1 + (1 | individual),
  beta ~ 1,
  nl = TRUE
)
fit <- brm(
  bform,
  data = dat
)
```

```{r}
bform <- bf(
  y ~ 0 + alpha + beta * (yt1 - alpha),
  alpha ~ 1 + (1 | individual),
  beta ~ 1 + (1 | individual),
  nl = TRUE
)
fit <- brm(
  bform,
  data = dat,
  file = "models/brm-example.rds"
)
```

```{r}
post_brms <- as_draws_df(fit) |> 
  select(
    alpha_hat = b_alpha_Intercept,
    alpha_scale = sd_individual__alpha_Intercept,
    beta_hat = b_beta_Intercept,
    beta_scale = sd_individual__beta_Intercept,
    sigma
  )

bind_rows(
  "Stan" = post_cmdstanr |> 
  pivot_longer(everything()),
  "brms" = post_brms |> 
  pivot_longer(everything()),
  .id = "Model"
) |> 
  mutate(side = if_else(Model == "Stan", "bottom", "top")) |> 
  ggplot(aes(value, name)) +
  scale_x_continuous(
    "Parameter value"
  ) +
  scale_color_brewer(
    palette = "Dark2",
    aesthetics = c("color", "fill")
  ) +
  stat_histinterval(
    aes(side = side, fill = Model, color = Model),
    normalize = "xy",
    scale = 0.5, 
    size = 1,
    breaks = 100,
    slab_alpha = 0.5
  ) +
  geom_point(
    data = as_tibble(dat) |> 
      select(sigma:beta_scale) |> 
      distinct() |> 
      pivot_longer(everything()),
    aes(color = "Real value", fill = "Real value"),
    size = 2
  ) +
  theme_linedraw() +
  theme(
    axis.title.y = element_blank()
  ) +
  labs(
    title = str_glue(
      "bf(y ~ 0 + alpha + beta * (yt1 - alpha),
          alpha ~ 1 + (1 | individual),
          beta ~ 1 + (1 | individual),
          nl = TRUE)"
    )
  ) +
  facet_wrap("name", scales = "free", ncol = 2)
```

# Next

Next problem is to generalize this to the multivariate situation where we have some (e.g.) X[t-1] predicting Y[t]. to be continued...
